Charity During Crackdown: Analyzing the Impact of State Repression of NGOs on Philanthropy

### 2020-01-20 manuscript/manuscript.Rmd
Some more edits

---
title: "Charity During Crackdown: Analyzing the Impact of State Repression of NGOs on Philanthropy"
short-title: Charity During Crackdown
author:
- name: Suparna Chaudhry
  affiliation: Christopher Newport University
  email: suparna.chaudhry@cnu.edu
  url: http://www.suparnachaudhry.com/
  twitter: SuparnaChaudhry
- name: Andrew Heiss
  affiliation: Brigham Young University
  email: andrew_heiss@byu.edu
  url: https://www.andrewheiss.com/
  twitter: andrewheiss
date: "`r format(Sys.time(), '%B %e, %Y')`"
git-repo: https://github.com/andrewheiss/ngo-crackdowns-philanthropy
reference-section-title: References
toc: false
title-page: true
abstract: >-
  State crackdown on non-governmental organizations (NGOs), particularly through legal restrictions on funding and advocacy, has become increasingly pervasive across the globe. How do legal restrictions on NGOs impact patterns of private philanthropic giving and individual donor decision-making? With reduced funding for NGOs working on contentious issues, and an absence of philanthropic culture in various developing countries, many NGOs rely on funds from private donors in Western countries. Existing research has predominantly focused on individual-level determinants of philanthropy. However, we argue that structural factors such as domestic political environments of NGO host countries can also influence preferences of foreign private donors, and that crackdown on NGOs serve as a heuristic to donors of organizational deservingness. Using a survey experiment, we explore how this crackdown changes donors' preferences based on the issue area and funding sources of the NGO. We find that while crackdowns do not directly influence the likelihood of donation on their own, donors are willing to donate substantially more to legally besieged NGOs abroad. We conclude by discussing the implications of our results on the sustainability of civil society organizations working in repressive countries. 
thanks: >-
  We thank Mark Buntaine, Michael DeCrescenzo, Jeffrey Friedman, Thomas Leeper, and Tristan Mahr for their helpful comments and input. We also thank Samantha Camilletti and Joshua Dutro for their excellent research assistance. This experiment received approval by the human subjects research committees at both Christopher Newport University (018.042) and Brigham Young University (E18104). We preregistered our hypotheses and research design at the Open Science Framework, and our preregistration protocol is available at <https://osf.io/dx973/>.
author-note: >-
  Suparna Chaudhry, PhD, Assistant Professor of Political Science, Christopher Newport University; Andrew Heiss, PhD, Visiting Assistant Professor, Romney Institute of Public Service and Ethics, Brigham Young University. 
additional-info: Figures should appear in color online only, not in print.
correspondence: >-
  Correspondence concerning this article should be addressed to Suparna Chaudhry, PhD, Department of Political Science, Christopher Newport University, Newport News, VA 23606, 757-594-7821. Contact: suparna.chaudhry@cnu.edu
word-count: "4,081"
keywords: international NGOs, donor behavior, philanthropy, human rights, repression
---

@@

# Load and clean data
likely <- readRDS(here("data", "derived_data", "results_models_likely.rds")) %>% 
  # Make LaTeXy versions of numbers
  mutate(delta = glue("$\\Delta = {x}$", x = tex_percent(theta_diff)),
         p_neq0 = glue("$p(\\Delta \\neq 0) = {x}$", x = round(p.diff.not0, 2))) %>% 
  # Simplify title
  mutate(title = str_replace_all(title, " \\| Crackdown", "")) %>% 
  # Add subscripts to funding and issue numbers
  mutate_at(vars(delta, p_neq0),
            funs(ifelse(title == "Funding",
                        str_replace(., "Delta", "Delta_\\\\text{funding}"), .))) %>% 
  mutate_at(vars(delta, p_neq0),
            funs(ifelse(title == "Issue",
                        str_replace(., "Delta", "Delta_\\\\text{issue}"), .)))

amount <- readRDS(here("data", "derived_data", "results_models_amount.rds")) %>% 
  # Make LaTeXy versions of numbers
  mutate(delta = glue("$\\Delta = {x}$", x = round(mu_diff, 2)),
         p_neq0 = glue("$p(\\Delta \\neq 0) = {x}$", x = round(p.diff.not0, 2))) %>% 
  # Simplify title
  mutate(title = str_replace_all(title, " \\| Crackdown", "")) %>% 
  # Add subscripts to funding and issue numbers
  mutate_at(vars(delta, p_neq0),
            funs(ifelse(title == "Funding",
                        str_replace(., "Delta", "Delta_\\\\text{funding}"), .))) %>% 
  mutate_at(vars(delta, p_neq0),
            funs(ifelse(title == "Issue",
                        str_replace(., "Delta", "Delta_\\\\text{issue}"), .)))

@@

Given increased crackdown on NGOs across the world, how do individual donors in the U.S. feel about donating to legally besieged NGOs abroad? Do legal restrictions on NGOs influence donors' decision to donate? Given information asymmetry and time constraints, individual donors often use heurestics to simply their decision-making [@TverskyKahneman:1974]. While much of this research has focused on heuristics from social networks and organizational practices [@CrosonShang:2011; @Tremblay-BoirePrakash:2017], in this research note, we argue that domestic political environments of NGO host countries can also serve as one such structural heuristic that can influence individual donor decisions. In particular, legal crackdowns on NGOs serve as a heuristic to donors of organizational deservingness and when donors are provided salient information about crackdown on NGOs and its impact on NGO activities, it can influence donor preferences and behavior. We use a survey experiment to measure the effect of crackdowns on stated donor behavior and find that while crackdowns do not directly influence the likelihood of donation on their own, respondents indicated willingness to donate substantially more to legally restricted NGOs. This effect persists when mediated by other organizational-level heuristics like NGO issue area and its main funding sources.

This research note two important contributions. While recent work has looked at how state repression of NGOs affects official aid [@ChaudhryHeiss:2018; @DupuyPrakash:2018], there is little research on how repression affects private aid. Further, while there is a robust existing literature on the individual-level determinants of philanthropy, far less is known about structural determinants that influence philanthropy, particularly towards international causes, and how these factors interact with organizational attributes to signal NGO deservingness.

@@

Substantial research on charitable-giving has looked at the motivations of private donors, with much of this work focusing on the individual-level determinants of giving [@Wiepking:2010; @BekkersWiepking:2011], and more recently, the effect of social networks on philanthropy [@HerzogYang:2018]. However, lesser research on focused on the impact of structural factors on the choice to donate. Moreover, research on individual donor behavior has overwhelming looked at giving to local, rather than international causes. While most research on giving to international causes has been restricted to elite, high net-worth donors [@USTrust2014], we explore individual giving to such causes becauses on individuals may rely on signals or heurestics that differ from the benchmarks that large donors (whether official or private) rely on.

The decision to donate to an NGO is rarely rational. In theory, perfectly rational donors should scrutinize each aspect of an NGO's organizational structure and programmatic performance prior to donation, but this rarely happens in practice [@Tremblay-BoirePrakash:2017]. Existing research shows that information on the efficiency of NGOs merely steers rather than encourage or discourage overall donations [@ryazanov:efficiency]. Similarly, the effects of changes in charity watchdog rations and nonprofit accountability systems on individual donors is also mixed - showing that donors do not have the time to research charities thoroughly [@Bekkers:2010; @SzperPrakash:2011]. Instead, donors respond to a host of heurestics and signals when deciding whether to donate to an NGOs, since seeking complete information about an organization's deservingness and efficiency is costly and time-consuming. These heuristics are commonly rooted in donor perceptions of organizational characteristics—donors make cursory judgments about an organization's issue area, mission, vision, and values, and seek out supplementary information from friends, family, and acquaintances [@SzperPrakash:2011; @Sloan:2009].

We argue that in addition to the above factors that signal an NGO's trustworthiness and deservingness, structural factors—including the nationality of the targets of NGO programs and whether an organization works abroad—can also serve as important heuristics in the decision to engage in philanthropy [@CasaleBaumann:2015; @KnowlesSullivan:2017; @Tremblay-BoirePrakash:2017]. We propose that the regulatory relationship between international NGOs and their host governments serves as a signal of NGO trustworthiness and deservingness. NGO legislation is not inherently restrictive, and governments frequently use regulations to benignly routinize the behavior of NGOs [@BloodgoodTB]. Restrictive NGO laws, on the other hand, are designed to limit organizational programming by imposing barriers to entry, funding, and advocacy. These stricter anti-NGO laws can signal to donors that governments perceive these groups as threatening and may crack down on them.

We argue that donors may process legal crackdowns differently from organizational-level heuristics - such as NGO missions, visions, values- because donors may be able to influence these latter attributes, at least to some extent. On the other hand, donors may perceive that they can do little to change the domestic political environment of another country. Such crackdowns can signal to donors that the government may eventually target other NGOs as well, which can increase donor urgency to support legally beseiged NGOs, even if donors are not completely satisfied with some organizational-level features of a group. As such, we expect that government restrictions abroad will increase respondents' likelihood of donating to the organization and increase the amount donated:

@@

@fig:causal-path represents this relationship more formally. Legal crackdowns (**C**) have a direct effect on donor outcomes (**Y**; donation likelihood and amount) because of their heuristic function. Legal crackdowns do not occur in isolation, however. Countries are more likely to repress NGOs when organizational missions (**I**) are in tension in with government preferences and when organizations receive substantial funding (**F**) from abroad [@DPR:2015]. Both of these organizational characteristics increase the likelihood of crackdown but also simultaneously serve as donor heuristics and influence perceptions of NGO deservingness.

@@

Another heurestic that individual donors may use for determining NGO deservingness is the *contentiousness of NGO issue areas*—or the degree to which an NGO's programming is compatible with government preferences.[^targeting] Organizations that address humanitarian issues such as relief and development have broader appeal to donors and these issues rarely challenge the government. Further, their programmatic output is easily quantifiable and donors can see the results of their charity more readily [@Bush:2015]. NGOs working on these issues can also frame their activities as in need of greater support. Perception of need, whether through personal experiences or media framing strongly influence individual decisions to donate to relief efforts [@BekkersWiepking:2010, 6-8].

@@

In these three sets of hypotheses, we do not assume perfect knowledge among individual donors. Donors may not always know about anti-NGO laws in target countries, or who funds an NGO, when they make their contributions. Rather, the three heuristics we explore here are potential signals to donors and our intention is to vary these three factors to test how powerful these factors are in comparison to other individual and organizational-level attributes. This can ultimately help NGOs optimize their fundraising strategies and tailor their fundraising attempts to private individuals.

@@

# Discussion: Does NGO crackdown affect individual philanthropy?

Dozens of countries have sought to restrict NGO funding and programming. Between the lack of a local philanthropic culture, an increasingly authoritarian approach that restricts space for domestic philanthropy, and the withdrawal of official governmental aid, many international NGOs have turned to foreign private donors as an additional funding source. Since seeking out perfect information about a nonprofit and its activities, especially internationally, is costly and time-consuming, donors use heuristics when deciding whom to support. Existing research has shown that donors take cues and heuristis from social netowrks, but can other information cues can influence donor preferences? This research note argues that domestic political environments of NGO host countries can serve as one such heuristic. Our results show that crackdowns provide donors with an additional signal of NGO deservingness and efficiency and can change both the likelihood that donors will give and the amount they will give. We find that while crackdowns do not directly influence the likelihood of donation on their own, respondents indicated willingness to donate `r percent_0(filter(amount, title == "Crackdown")$pct_change)` more to legally restricted NGOs. This effect persists when mediated by other heuristics, with respondents giving `r percent_0(filter(amount, title == "Humanitarian assistance")$pct_change)` more to humanitarian NGOs, and `r percent_0(filter(amount, title == "Human rights | Private")$pct_change)` more to privately-funded human rights NGOs. Anti-NGO crackdowns can thus serve as a signal of NGO effectiveness.

This research note should be viewed as an exploratory study, and it raises numerous quesitions for future research. These include determining the extent to which these preferences translate into behavior—here, we have only asked respondents to make a hypothetical donation, but preferences may not reflect actual behavior. Future research can also test the generalizability of our findings to other issue areas.

@@

[^targeting]: While anti-NGO legislation (in the form of barriers to entry, funding, and advocacy) can theoretically tatget all NGOs (or a large subset of them) within a state's borders, in practices, these laws are enforced more selectively against contentious groups that pose threats to states. See Chaudhry (2016) and Heiss (2017).

>>>>

---
title: "Charity During Crackdown: Analyzing the Impact of State Repression of NGOs on Philanthropy"
short-title: Charity During Crackdown
author:
- name: Suparna Chaudhry
  affiliation: Christopher Newport University
  email: suparna.chaudhry@cnu.edu
  url: http://www.suparnachaudhry.com/
  twitter: SuparnaChaudhry
- name: Andrew Heiss
  affiliation: Brigham Young University
  email: andrew_heiss@byu.edu
  url: https://www.andrewheiss.com/
  twitter: andrewheiss
date: "`r format(Sys.time(), '%B %e, %Y')`"
git-repo: https://github.com/andrewheiss/ngo-crackdowns-philanthropy
reference-section-title: References
toc: false
title-page: true
abstract: >-
  State crackdown on non-governmental organizations (NGOs), particularly through legal restrictions on funding and advocacy, has become increasingly pervasive across the globe. How do legal restrictions on NGOs impact patterns of private philanthropic giving and individual donor decision-making? With reduced funding for NGOs working on contentious issues, and an absence of philanthropic culture in various developing countries, many NGOs rely on funds from private donors in Western countries. Existing research has predominantly focused on individual-level determinants of philanthropy. However, we argue that structural factors such as domestic political environments of NGO host countries can also influence preferences of foreign private donors, and that crackdown on NGOs serve as a heuristic to donors of organizational deservingness. Using a survey experiment, we explore how this crackdown changes donors' preferences based on the issue area and funding sources of the NGO. We find that while crackdowns do not directly influence the likelihood of donation on their own, donors are willing to donate substantially more to legally besieged NGOs abroad. We conclude by discussing the implications of our results on the sustainability of civil society organizations working in repressive countries. 
thanks: >-
  We thank Mark Buntaine, Michael DeCrescenzo, Jeffrey Friedman, Thomas Leeper, and Tristan Mahr for their helpful comments and input. We also thank Samantha Camilletti and Joshua Dutro for their excellent research assistance. This experiment received approval by the human subjects research committees at both Christopher Newport University (018.042) and Brigham Young University (E18104). We preregistered our hypotheses and research design at the Open Science Framework, and our preregistration protocol is available at <https://osf.io/dx973/>.
author-note: >-
  Suparna Chaudhry, PhD, Assistant Professor of Political Science, Christopher Newport University; Andrew Heiss, PhD, Assistant Professor of Public Management, Andrew Young School of Policy Studies, Georgia State University. 
additional-info: Figures should appear in color online only, not in print.
correspondence: >-
  Correspondence concerning this article should be addressed to Suparna Chaudhry, PhD, Department of Political Science, Christopher Newport University, Newport News, VA 23606, 757-594-7821. Contact: suparna.chaudhry@cnu.edu
word-count: "4,081"
keywords: international NGOs, donor behavior, philanthropy, human rights, repression
---

@@

# Load and clean data
likely <- readRDS(here("data", "derived_data", "results_models_likely.rds")) %>%
  # Make LaTeXy versions of numbers
  mutate(delta = glue("$\\Delta = {x}$", x = tex_percent(theta_diff)),
         p_neq0 = glue("$p(\\Delta \\neq 0) = {x}$", x = round(p.diff.not0, 2))) %>%
  # Simplify title
  mutate(title = str_replace_all(title, " \\| Crackdown", "")) %>%
  # Add subscripts to funding and issue numbers
  mutate_at(vars(delta, p_neq0),
            list(~ifelse(title == "Funding",
                         str_replace(., "Delta", "Delta_\\\\text{funding}"), .))) %>%
  mutate_at(vars(delta, p_neq0),
            list(~ifelse(title == "Issue",
                         str_replace(., "Delta", "Delta_\\\\text{issue}"), .)))

amount <- readRDS(here("data", "derived_data", "results_models_amount.rds")) %>%
  # Make LaTeXy versions of numbers
  mutate(delta = glue("$\\Delta = {x}$", x = round(mu_diff, 2)),
         p_neq0 = glue("$p(\\Delta \\neq 0) = {x}$", x = round(p.diff.not0, 2))) %>%
  # Simplify title
  mutate(title = str_replace_all(title, " \\| Crackdown", "")) %>%
  # Add subscripts to funding and issue numbers
  mutate_at(vars(delta, p_neq0),
            list(~ifelse(title == "Funding",
                         str_replace(., "Delta", "Delta_\\\\text{funding}"), .))) %>%
  mutate_at(vars(delta, p_neq0),
            list(~ifelse(title == "Issue",
                         str_replace(., "Delta", "Delta_\\\\text{issue}"), .)))

@@

Given increased crackdown on NGOs across the world, how do individual donors in the U.S. feel about donating to legally besieged NGOs abroad? Do legal restrictions on NGOs influence donors' decision to donate? Given information asymmetry and time constraints, individual donors often use heuristics to simply their decision-making [@TverskyKahneman:1974]. While much of this research has focused on heuristics from social networks and organizational practices [@CrosonShang:2011; @Tremblay-BoirePrakash:2017], in this research note, we argue that domestic political environments of NGO host countries can also serve as one such structural heuristic that can influence individual donor decisions. In particular, legal crackdowns on NGOs serve as a heuristic to donors of organizational deservingness and when donors are provided salient information about crackdown on NGOs and its impact on NGO activities, it can influence donor preferences and behavior. We use a survey experiment to measure the effect of crackdowns on stated donor behavior and find that while crackdowns do not directly influence the likelihood of donation on their own, respondents indicated willingness to donate substantially more to legally restricted NGOs. This effect persists when mediated by other organizational-level heuristics like NGO issue area and its main funding sources.

This research note makes two important contributions. While recent work has looked at how state repression of NGOs affects official aid [@ChaudhryHeiss:2018; @DupuyPrakash:2018], there is little research on how repression affects private aid. Further, while there is a robust existing literature on the individual-level determinants of philanthropy, far less is known about structural determinants that influence philanthropy, particularly towards international causes, and how these factors interact with organizational attributes to signal NGO deservingness.

@@

Substantial research on charitable-giving has looked at the motivations of private donors, with much of this work focusing on the individual-level determinants of giving [@Wiepking:2010; @BekkersWiepking:2011], and more recently, the effect of social networks on philanthropy [@HerzogYang:2018]. However, little research has focused on the impact of structural factors on the choice to donate. Moreover, research on individual donor behavior has overwhelming looked at giving to local, rather than international causes. While most research on giving to international causes has been restricted to elite, high net-worth donors [@USTrust2014], we explore private philanthropic giving because individuals rely on signals and heuristics that differ from the benchmarks that large donors (whether official or private) rely on.

The decision to donate to an NGO is rarely rational. In theory, perfectly rational donors should scrutinize each aspect of an NGO's organizational structure and programmatic performance prior to donation, but this rarely happens in practice [@Tremblay-BoirePrakash:2017]. Existing research shows that information on the efficiency of NGOs merely steers rather than encourage or discourage overall donations [@ryazanov:efficiency]. Similarly, the effects of changes in charity watchdog ratings and nonprofit accountability systems on individual donors is also mixed, showing that donors do not have the time to research charities thoroughly [@Bekkers:2010; @SzperPrakash:2011]. Instead, donors respond to a host of heuristics and signals when deciding whether to donate to an NGOs, since seeking complete information about an organization's deservingness and efficiency is costly and time-consuming. These heuristics are commonly rooted in donor perceptions of organizational characteristics—donors make cursory judgments about an organization's issue area, mission, vision, and values, and seek out supplementary information from friends, family, and acquaintances [@SzperPrakash:2011; @Sloan:2009].

In addition to the above factors that signal an NGO's trustworthiness and deservingness, we argue that structural factors—including the nationality of the targets of NGO programs and whether an organization works abroad—can also serve as important heuristics in the decision to engage in philanthropy [@CasaleBaumann:2015; @KnowlesSullivan:2017; @Tremblay-BoirePrakash:2017]. We propose that the regulatory relationship between international NGOs and their host governments serves as a signal of NGO trustworthiness and deservingness. NGO legislation is not inherently restrictive, and governments frequently use permissive regulations to benignly routinize the behavior of NGOs [@BloodgoodTB; @DeMattee:2018]. Restrictive NGO laws, on the other hand, are designed to limit organizational programming by imposing barriers to entry, funding, and advocacy. These stricter anti-NGO laws can signal to donors that governments perceive these groups as threatening and may crack down on them.

We argue that donors perceive legal crackdowns differently from organizational-level heuristics—such as NGO missions, visions, values—because donors may be able to influence these latter attributes, at least to some extent. On the other hand, donors can do little to change the domestic political environment of another country. Legal crackdowns abroad can signal to donors that the government may eventually target other NGOs as well, which can increase donor urgency to support legally besieged NGOs, even if donors are not completely satisfied with some organizational-level features of a group. As such, we expect that government restrictions abroad will increase respondents' likelihood of donating to the organization and increase the amount donated:

@@

@fig:causal-path represents this relationship more formally. Legal crackdowns (**C**) have a direct effect on donor decisions (**Y**; donation likelihood and amount) because of their heuristic function. Legal crackdowns do not occur in isolation, however. Countries are more likely to repress NGOs when organizational missions (**I**) are in tension in with government preferences and when organizations receive substantial funding (**F**) from abroad [@DPR:2015]. Both of these organizational characteristics increase the likelihood of crackdown but also simultaneously serve as donor heuristics and influence perceptions of NGO deservingness.

@@

Another heuristic that individual donors may use for determining NGO deservingness is the *contentiousness of NGO issue areas*—or the degree to which an NGO's programming is compatible with government preferences.[^targeting] Organizations that address humanitarian issues such as relief and development have broader appeal to donors and these issues rarely challenge the government. Further, their programmatic output is easily quantifiable and donors can see the results of their charity more readily [@Bush:2015]. NGOs working on these issues can also frame their activities as in need of greater support. Perception of need, whether through personal experiences or media framing strongly influence individual decisions to donate to relief efforts [@BekkersWiepking:2010, 6-8].

@@

Importantly, when testing these hypotheses, we do not assume that individual donors have perfect knowledge about an NGO or the countries an NGO works in. Donors may not always know about anti-NGO laws in target countries or who funds an NGO when they make their contributions. Rather, the heuristics we explore here are shortcuts to donor information-gathering. To measure the effect of these signals on the propensity and amount of individual donations, we use an experiment to vary the different organizational and structural signals donors are exposed to. Comparing these effects allows us to measure the relative strength of these heuristics, and ultimately help NGOs optimize and tailor their fundraising strategies for private individuals.

@@

# Discussion: Do crackdowns on NGOs affect individual philanthropy?

Dozens of countries have sought to restrict NGO funding and programming. Between the lack of a local philanthropic culture, an increasingly authoritarian approach that restricts space for domestic philanthropy, and the withdrawal of official governmental aid, many international NGOs have turned to foreign private donors as an additional funding source. Since seeking out perfect information about a nonprofit and its activities, especially internationally, is costly and time-consuming, donors use heuristics when deciding whom to support. Existing research has shown that donors take cues and heuristics from social networks, but can other information cues influence donor preferences? We argue that domestic political environments of NGO host countries can indeed serve as one such heuristic. Our results show that crackdowns provide donors with an additional signal of NGO deservingness and efficiency and can change both the likelihood that donors will give and the amount they will give. We find that while crackdowns do not directly influence the likelihood of donation on their own, respondents indicated willingness to donate `r percent_0(filter(amount, title == "Crackdown")$pct_change)` more to legally restricted NGOs. This effect persists when mediated by other heuristics, with respondents giving `r percent_0(filter(amount, title == "Humanitarian assistance")$pct_change)` more to humanitarian NGOs, and `r percent_0(filter(amount, title == "Human rights | Private")$pct_change)` more to privately-funded human rights NGOs. Anti-NGO crackdowns can thus serve as a signal of NGO effectiveness.

This research note should be viewed as an exploratory study, and it raises numerous questions for future research. These include determining the extent to which these preferences translate into behavior—here, we have only asked respondents to make a hypothetical donation, but preferences may not reflect actual behavior. Future research can also test the generalizability of our findings to other issue areas. TODO OTHER THINGS: Crackdown in authoritarian regimes, but there's an increase in restrictive laws in democracies [@Swiney:2019]

@@

[^targeting]: While anti-NGO legislation (in the form of barriers to entry, funding, and advocacy) can theoretically target all NGOs (or a large subset of them) within a state's borders, in practices, these laws are enforced more selectively against contentious groups that pose threats to states. See @Chaudhry:2016 and @Heiss:2017.
